Fractional Types

### 2019-11-18 fracGC/p.tex
paper

\paragraph*{Quantum Error Correction.} The main requirement for
quantum error correction is the ability to replicate the state of a
qubit onto a number of qubits~\cite[Ch.~3]{NAP25196}. This is however
impossible to do directly by the no-cloning theorem. Again ancilla
qubits come to the rescue: because ancilla qubits start in a known
initial state, it is possible to create a simple circuit that makes
their output state match a protected qubit without knowing or
disturbing the protected qubit. Chong et al.~\cite{sigarchblog}
illustrate the idea with a simple example:
\begin{quote}
consider a 3-qubit quantum majority code in which a logical ``0'' is
encoded as ``000'' and a logical ``1'' is encoded as ``111.''  Just as with
a classical majority code, a single bit-flip error can be corrected by
restoring to the majority value.  Unlike a classical code, however, we
can not directly measure the qubits else their quantum state will be
destroyed.  Instead, we measure syndromes from each possible pair of
qubits by interacting them with an ancilla, then measure each ancilla.
Although the errors to the qubits are actually continuous, the effect
of measuring the ancilla is to discretize the errors, as well as
inform us whether an error occurred so that it can be corrected.  
\end{quote}

@@

The step from reversible classical circuits to quantum circuits
requires the addition of just one gate that creates
superpositions. Similarly the step from a reversible classical
programming language to a quantum one requires just the management of
superpositions~\cite{10.1007/978-3-319-89366-2_19}. In the cited
paper, the core reversible classical language was Theseus~\cite{XXX},
a language that adds convenient syntactic extensions over an more
basic language $\Pi$ of combinators witnessing isomorphisms over
finite types~\cite{XXX}. The combinator-based language has the
advantage of being convenient for formal analysis for at least two
reasons: (i) it is conceptually simple and small, and (ii) it has
direct and evident connections to type theory and category
theory. Indeed our solution for managing ancillae is inspired by the
construction of \emph{compact closed categories}~\cite{XXX}. These
categories extend the monoidal categories~\cite{XXX} which are to
model most models of resource-aware programming languages~\cite{XXX}
with a new type constructor that creates duals or inverses to existing
types. This will be our fractional type.

@@

@@

@@

>>>>

\paragraph*{Quantum Error Correction.} The main requirement for
quantum error correction is the ability to replicate the state of a
qubit onto a number of qubits~\cite[Ch.~3]{NAP25196}. This is however
impossible to do directly by the no-cloning theorem~\cite{XXX}. Again ancilla
qubits come to the rescue: because ancilla qubits start in a known
initial state, it is possible to create a simple circuit that makes
their output state match a protected qubit without knowing or
disturbing the protected qubit. Chong et al.~\cite{sigarchblog}
illustrate the idea with a simple example:
\begin{quote}
consider a 3-qubit quantum majority code in which a logical ``0'' is
encoded as ``000'' and a logical ``1'' is encoded as ``111.''  Just as with
a classical majority code, a single bit-flip error can be corrected by
restoring to the majority value.  Unlike a classical code, however, we
can not directly measure the qubits else their quantum state will be
destroyed.  Instead, we measure syndromes from each possible pair of
qubits by interacting them with an ancilla, then measure each ancilla.
Although the errors to the qubits are actually continuous, the effect
of measuring the ancilla is to discretize the errors, as well as
inform us whether an error occurred so that it can be corrected.  
\end{quote}

@@

The step from reversible classical circuits to quantum circuits
requires the addition of just one gate that creates
superpositions. Similarly the step from a reversible classical
programming language to a quantum one requires just the management of
superpositions, for example starting from the classical reversible
language Theseus~\cite{XXX}, a small extension produces a quantum
language with quantum control~\cite{10.1007/978-3-319-89366-2_19}. The
language Theseus~\cite{XXX} mentioned above is a language that uses
convenient syntactic extensions over an more basic language $\Pi$ of
combinators witnessing isomorphisms over finite types~\cite{XXX}. The
combinator-based language has the advantage of being more amenable to
formal analysis for at least two reasons: (i) it is conceptually
simple and small, and (ii) it has direct and evident connections to
type theory and category theory. Indeed our solution for managing
ancillae is inspired by the construction of \emph{compact closed
  categories}~\cite{XXX}. These categories extend the monoidal
categories~\cite{XXX} which are used to model many resource-aware
programming languages~\cite{XXX} (including $\Pi$) with a new type
constructor that creates duals or inverses to existing types. This
dual will be our fractional type.

@@

move from 1/t to 1/v (how the quantum search runs; need ancilla 
initialized to a specific value)

@@

PiPointedFrac:
lift pi
lift multiplicative structure
monad/comonad for singletons
eta/epsilon

rationale: we want singleton types to track, in the type system
the values of ancillae, i.e.. we construct a proof using the type
system. Technically the singleton types allow you to focus on one
value by reflecting it in the type system. Now this type can only be
equal to another type if and only if the other type is the same
singleton value.

Once singleton types are built, we need to make sure circuit
evaluation is also reflected in the type system. So in fact
computation itself is reflected in the type system. In order to mix
singleton types with other types, we use pointed types as a unifying
substrate. Now all types have values; we can focus (using monad) and
unfocus using (comonad). (This by the way ensures we don't write 1/0
which woudl be bad).

@@

allocate A
allocate B
combinate the gc's for A and B to get 1/(AxB)
allocate AxB: split the gc in two 1/A and 1/B
use the different gcs to satisfy each other
